Supporting Information\\ \textit{Cross Validation Approaches for Penalized Cox Regression}

### 2020-05-06 supplementary.tex
Edits to deviance residuals (section and supplement)

\section{Baseline Estimation for Cross-Validated Deviance Residuals}
\subsection{Cross-Validated Deviance Residuals}
The Cox model implies the following relationship between the baseline cumulative hazard and the cumulative hazard for individual $i$:
\begin{equation}
  \hat{\Lambda}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta).
\end{equation}
Consider the following approach: for fold $k$, we obtain $\hat{\beta}^{-k}$ from the training set $T_k$ and then for each $i$ in the test set $D_k$, 
\begin{equation}
 \label{eq:cv-bhz}
  \hat{\Lambda}^{cv}_{i}(t) =  \hat{\Lambda}_{0}(t)\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}

A natural candidate for incorporating the cumulative hazard into a loss function is the deviance residual, a normalized form of the Martingale residual \citep{Therneau1990}.  Given \eqref{eq:cv-bhz}, we can first obtain the cross-validated Martingale residual: 
\begin{equation}
  \hat{M^{cv}_{i}} = \delta_{i} - \hat{\Lambda}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}).
\end{equation}
The cross-validated deviance residuals can then be derived from the Martingale residuals: 
\begin{equation} 
  d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}.
\end{equation}
The sum of squared cross-validated deviance residuals, $\sum_{i}\hat{d}_{i}^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression. We refer to this method as the \emph{cross-validated deviance residuals} approach.

@@

	\subsubsection{Nelson-Aalen Estimator}
	 The simplest approach to estimating the cumulative baseline hazard, $\Lambda_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{\Lambda}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
	The Nelson-Aalen Estimator does not involve any information from the covariates. It only depends on the failure times.

@@

	\subsubsection{Kalbfleisch and Prentice's Approach}					
	\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates. Assuming only a single failure occurs at $t_i$, based on the fitted values of $\hat{\bbeta}$, we can first compute linear predictors $\hat{\eta}_i$ at each observed failure time. Based on $\hat{\eta}_i$, we can compute a discrete hazard component at each observed failure time):
\begin{equation}
1 - \hat{\alpha}_{i} = 1 - \left\{ 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}exp (\hat{\eta}_{j}} \right\}^{exp(-\hat{\eta}_{i})},
\end{equation}
and aggregate them together into the cumulative hazard function:
\begin{equation} 
\hat{\Lambda}_0(t) = \sum(1 - \hat{\alpha}_i)\mathds{1}(t_{i} \leq t),
\end{equation}

>>>>

\section{Baseline Estimation for Cross-Validated Deviance Residuals}

As described in the manuscript, the Martingale residuals $\{m_i\}_{i=1}^n$ and deviance residuals $\{d_i\}_{i=1}^n$ for the Cox model are given by
\begin{align*}
  m_{i} &= \delta_{i} - H_{0}(t_{i})\exp(X_i\Tr\beta) \\
  d_{i} &= \text{sign}(m_{i})\sqrt{-2(m_i + \delta_{i}\log(\delta_{i} - m_i))},
\end{align*}
where $H_0$ is the cumulative baseline hazard.  Our cross-validated versions of these residuals use
\begin{align*}
  m_{i}^{cv} &= \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k}) \\
  d_{i}^{cv} &= \text{sign}(m_{i}^{cv})\sqrt{-2(m_i^{cv} + \delta_{i}\log(\delta_{i} - m_i^{cv}))},
\end{align*}
for $i$ in the test set $D_k$, with $\hat{\beta}^{-k}$ obtained from the training set $T_k$.  The sum of squared cross-validated deviance residuals, $\sum_{i}(d_{i}^{cv})^2$, are then used as the cross validated error, analogously to using the residual sum of squares in linear regression.  In the manuscript, this method is referred to as the \emph{cross-validated deviance residuals} approach.

@@

	\subsubsection{Nelson-Aalen Estimator}
	 The simplest approach to estimating the cumulative baseline hazard, $H_0$, is the Nelson-Aalen Estimator \citep{nelson1969, aalen1978}:
\begin{equation}
  \hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}.
\end{equation}
Under the assumption that there is no ties among failure times, the numerator in the summation is either 0 or 1. The Nelson-Aalen Estimator depends only on the failure times and does not involve any information from the covariates. Procedure~\ref{alg:na} provides an explicit description of this algorithm.

\begin{algorithm}
\SetAlgoLined
Use all observations' time and status to compute $\hat{H}_{0}(t) = \sum_{t_j \leq t}\frac{\text{number of failures at time } t_j}{\text{number at risk right before time }t_j}$\;
\For{each regularization parameter $\lambda$}{
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{\label{alg:na} Compute cross-validated deviance residuals based on the Nelson-Aalen estimator}
\end{algorithm}

@@

In the original proposal of the Martingale Residuals \citep{Therneau1990}, Breslow's estimator \citep{Breslow1974} was the prefered approach for estimating the cumulative baseline hazard. Under the assumption that there is no ties among the failure times, the Breslow's estimator is defined as:
\begin{equation} 
\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}.
\end{equation}
Covariates of the fitted model are incorporated in the denominator of the summation.
Procedure~\ref{alg:br} provides an explicit description of this algorithm.
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t}\frac{\delta(t_i)}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})}$\;
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{\label{alg:br} Compute cross-validated deviance residuals based on the Breslow estimator}
\end{algorithm}

\subsubsection{Kalbfleisch and Prentice's Approach}

\cite{Kalbfleisch2011} proposed a cumulative hazard estimator that incorporates covariates. Assuming only a single failure occurs at $t_i$, based on the fitted values of $\hat{\bbeta}$, we can first compute linear predictors $\hat{\eta}_i$ at each observed failure time. Based on $\hat{\eta}_i$, we can compute a discrete hazard component at each observed failure time):
\begin{equation}
1 - \hat{\alpha}_{i} = 1 - \left( 1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right)^{\exp(-\hat{\eta}_{i})},
\end{equation}
and aggregate them together into the cumulative hazard function:
\begin{equation} 
\hat{H}_0(t) = \sum_{t_{i} \leq t}(1 - \hat{\alpha}_i),
\end{equation}
Procedure~\ref{alg:kp} provides an explicit description of this algorithm.
\begin{algorithm}
\SetAlgoLined
\For{each regularization parameter $\lambda$}{
 Fit the model using all observations' time and status \;
 Obtain linear predictor $\hat{\eta}$ from the fitted model\;
 Compute $\hat{H}_0(t) = \sum_{t_{i} \leq t} \left\{1 -  \left(1 - \frac{\exp (\hat{\eta}_{i})}{\sum_{ j \in R(t_{i})}\exp (\hat{\eta}_{j})} \right) ^{\exp(-\hat{\eta}_{i})}\right\}$\;
 \For{each fold $k = 1,2,...,K$}{
  Use the training set $T_k$ to obtain $\hat{\beta}^{-k}$\;
  	\For{each obervation $i$ in the $k$th fold}{
  	Compute the Cross-validated Martingale Residual: $\hat{M}^{cv}_{i}  = \delta_{i} - \hat{H}_{0}(t_{i})\exp(X_i\Tr\hat\beta^{-k})$\;
  	Compute the Deviance Residual: $ d_{i} = \text{sgn}(\hat{M}^{cv}_{i})\sqrt{-2(\hat{M}^{cv}_{i} + \delta_{i}\log(\delta_{i} - \hat{M}^{cv}_{i}))}$\;
  	}
 }
 Compute the sum of squares: $\sum_{i}\hat{d}_{i}^2$
 }
 \caption{\label{alg:kp} Compute cross-validated deviance residuals based on the Kalbfleisch and Prentice estimator}
\end{algorithm}
